Toward a novel piezoelectric energy harvester for earcanal dynamic motion exploitation, using a bistable resonator cycled by coupled hydraulic valves made of collapsed flexible tubes

### 2022-11-01 article_system.tex
Recorrection Fabien sur Overleaf...

\title{Toward a novel piezoelectric energy harvester for ear canal dynamic motion exploitation, using a bistable resonator cycled by coupled hydraulic valves made of collapsed flexible tubes}

@@

\author[critias]{Michel DEMUYNCK}
\author[critias]{Aidin DELNAVAZ}
\author[critias]{Jérémie VOIX}

\begin{abstract}
Scavenging energy from the ear canal dynamic motion during jaw motion may be a practical way to enhance the battery autonomy of hearing aids. One of the main challenges is to optimize the extracted energy while coping with the constraints associated to soft human tissues and restricted ear canal volume. We propose a new concept of harvester made of a liquid filled earplug transferring the energy outside the ear canal to a generator. It is composed oa hydraulic amplifier, two hydraulic cylinders actuating a bistable resonator raising the source frequency driving an amplified piezoelectric transducer to generate electricity. The cycling of the resonator is achieved using two innovative flexible hydraulic valves based on the buckling of flexible tubes.\\
A multiphysic coupled model is established in order to determine the system operation requirements and to evaluate its theoretical performance. It exhibits a theoretical energy conversion efficiency of 85\%. The electromechanical performance of the resonator couple to the piezoelectric transducer as well as the hydraulic behavior of the valves are experimentally investigated. The global model is updated with the experimental data to ameliorate its predictability for a further optimized design. The energy losses are identified, which allow an entire design proposed to improve the experimental energy conversion efficiency to 26\%.

\end{abstract}

\begin{keyword}
Energy harvesting, frequency-up, ear canal, multiphysic, hydraulic valve
\end{keyword}

@@

The growing use of wireless devices and the miniaturization of electronic circuits have led to significant progress on the energy consumption of mobile devices on the human body environment. Energy harvesting methods have been studied in order to complement their power supply and enhance the autonomy of the batteries. In-ear devices such as hearing aids and cochlear implants are powered by disposable cells or rechargeable batteries. The energy consumption of the best integrated devices in the literature rises up to $17$J for a 10 hours daily use \cite{Scherer2019,Yip2015,Kulah2022}. Woodruff \emph{et al.} underlined that patients using hearing aids sometimes struggle to change and select batteries for their devices \cite{Woodruff2021}. Another statistical study revealed that the majority of hearing aids and cochlear implants users prefer disposable batteries for the long autonomy, but would have liked rechargeable solutions if the battery cycle was sufficient to complete the day \cite{PracticesAudiology2016}. Also, the long term use makes the rechargeable solutions a more economical and more ecological choice compared to the disposable cells. These arguments motivated the development of energy harvesting systems.

The common energy sources are solar or wind energy but their availability in the low size environment of the application context makes them not suitable for exploitation. The activities of the human body generate a considerable amount of energy of different natures. Among them, the electrochemical energy from the inner ear \cite{Mercier2012}, the kinetic energy from walking or head movements \cite{Azimi2021,Smilek2016}, the body heat thermal energy \cite{Kim2014}, the strain energy from the skin deformation \cite{Jin2021} could be harvested for in-ear applications. An interesting energy source for cochlear implants or hearing aids appears to be the ear canal mechanical deformation as it is directly located in the area of the application. Delnavaz \emph{et al.} first showed in 2012 that the jaw movements changes de ear canal geometry during mastication. Carioli \emph{et al.} then used a customized earplug molding technic with a 3D scanner to reveal a global bending movement and a local compression area in the ear canal during mastication \cite{Carioli2016}. These two mechanical deformations represent a potential energy source that can be exploited.

Delnavaz \emph{et al.} developed a flexible piezoelectric harvester made of a silicon earplug with a PVDF layer \cite{Delnavaz2013}. The earplug has been molded on the subject in order to maximize the contact surface with the ear canal. The harvester is capable of exploiting both the flexion and the compression energy sources. The experimental prototype was able to extract $44$µJ from a mastication cycle. The soft materials facilitate the bio-integration of the harvester but the PVDF low electromechanical coupling coefficient could have led to poor energy conversion efficiency. A hydro-electromagnetic energy harvester was also developed by the same researchers \cite{Delnavaz2012}. The energy is extracted by a liquid filled earplug whose internal volume vary as a result of the radial compression of the ear canal internal walls during the jaw movements. This variation put into motion a moving magnet translating in a water column surrounded by a coil. The generated energy experimentally amounts to 0.2µJ per mastication cycle. The prototype appeared to be less efficient than the PVDF earplug. This could be explained by the poor power generation of the electromagnetic transduction for a low frequency energy source (1.57Hz) and a small scale application \cite{Kulah2008,Priya2017}. However, the hydraulic transmission allows the use of more complex and voluminous technological solutions since the energy becomes available outside the small ear canal volume.

Based on the analysis of the literature, we propose a new solution optimizing the energy conversion efficiency to maximize the harvested energy from the ear canal dynamic motion. The next section states on the hydraulic energy source characterization and on the harvesting strategy that we adopted. Section \ref{sec:HARVESTER PRESENTATION AND OPERATION PRINCIPLE} presents the harvester and describes its operation principle. Section \ref{sec:SYSTEM MODELING} details the global system modeling and section \ref{sec:NUMERICAL MODEL AND SIMULATIONS} presents the numerical simulation of the theoretical behavior of the harvester. Then section \ref{sec:EXPERIMENTAL CHARACTERIZATIONS OF THE ELECTROMECHANICAL CONVERTER} then exposes the experimental validation performed on the electromechanical converter stage and section \ref{sec:EXPERIMENTAL APPROACH FOR THE HV DESIGN} shows the experimental approach for the design of the hydraulic valves for the system cycling. Then, section \ref{sec:MODEL RECALIBRATION WITH EXPERIMENTAL DATA} presents the experimentally adjusted global harvester model and discusses this work major contributions and the harvester potential. Finally, a conclusion is stated.

%/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\%
\section{ENERGY SOURCE AND HARVESTING STRATEGY}
\label{sec:THE ENERGY SOURCE AND HARVESTING STRATEGY}
%/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\%
    %///////////////////////////////////////////// 
	\subsection{The energy source}	
	\label{The energy source}
    %/////////////////////////////////////////////
Human tissue requires soft flexible materials in order to extract the ear canal deflection energy without causing discomfort. However, flexible transducers admit low energy conversion efficiency as a part of the extracted energy is dissipated in the elastic non-active material. Additionally, the low frequency of mastication makes it difficult to use resonant structures that generally admit better conversion efficiencies than low frequency solutions \cite{Ashraf2011}. Thus, the traditional transducing methods are not suitable enough for the application.

To date, the best solution proposed in the literature consists on using soft materials as the piezoelectric PVDF earplug previously introduced \cite{Delnavaz2013}. The low mechanical impedance of the soft tissue in fact limits the use of more effective rigid piezoelectric materials.

The hydraulic extraction strategy proposed in \cite{Delnavaz2012} is less effective than the piezoelectric earplug but allows to transmit the energy outside the ear canal and benefit so from larger effective space. The latter is in fact essential as it becomes then possible to implement an impedance adaption stage and more effective transducing methods that could not fit directly inside the ear canal restricted volume. In fact, Marin \emph{et al.} demonstrated that the power generated by piezoelectric or electromagnetic transducers is as much better as the effective volume of transducing material is high \cite{Marin2011}. The present work is then based on the exploitation of the ear canal by a liquid filled earplug. As energy source, it can be assimilated to a micro pump characterized by its volume variation $V_{ear}$ and pressure variation $p_{ear}$. In order to optimize the energy extraction, the earplug must fit the ear canal wall. Turcot \emph{et al.} demonstrated that the initial pressurization must reach a minimum of $14$kPa to achieve a good fit \cite{TURCOT2011}. Bouchard-Roy \emph{et al.} studied the dynamic pressure variation inside the earplug for different initial pressurization levels \cite{Bouchard-Roy2020}. A maximum of $\text{max}(\Delta p_{ear})=12$kPa of dynamic pressure variation, with a $28$kPa of initial pressurization, has been noted during the mastication for seven subjects. Besides, the temporal evolution of the volume variation $\Delta V_{ear}$ has been estimated in \cite{Delnavaz2012} and the result for a mastication cycle is presented on figure \ref{fig:deltaV_ear}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!htbp]
	\centering
	\captionsetup{justification=centering}
	\includegraphics[trim={20.5cm 0cm 0cm 10.8cm},clip, width=0.4\textwidth]{figures/deltaV_ear.pdf}
	\caption{Ear canal volume variation for one mastication cycle}
	\label{fig:deltaV_ear}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The earplug operation characteristic behavior has not been studied yet. Thus, the evolution of the volume variation $\Delta V_{ear}$ is not well known as a function of the pressure variation $\Delta p_{ear}$. Then, considering the pressure and volume variation levels from \cite{Delnavaz2012} and \cite{Bouchard-Roy2020}, the equation \ref{eq:max_hydraulic_energy} approximates an upper limit for the extractable hydraulic energy with the maximum values of $\Delta p_{ear}$ et $\Delta V_{ear}$ that have been recorded, i.e. $p_c=12kPa$ and $(\Delta V_{ear})_{max}=60$mm$^3$. $p_c$ stands for the comfort pressure by assuming that the user could feel discomfort if $\Delta p_{ear}>p_c$. The energy estimation presumes that the earplug is a perfect flow rate source, which is a theoretical hypothesis leading to $0.9$mJ available from one mastication. Based on 2200 mastication cycles per day, according to \cite{Goll2011}, there would be $2$mJ available per day, per ear. This amount of energy could theoretically enhance supply $23.5$\% of the $17$J of needed for the state of the art of cochlear implants \cite{Kulah2022}. Further research in both domains related to hearing aids and energy harvesting could lead to fully autonomous devices in the future.
\begin{align}
	\text{max}(E_{hyd}) = p_c \cdot (\Delta V_{ear})_{max} = 0.9\text{mJ}
	\label{eq:max_hydraulic_energy}
\end{align}

    %///////////////////////////////////////////// 
	\subsection{Harvesting strategy}	
	\label{The harvesting strategy}
    %/////////////////////////////////////////////
Electromagnetic transducers are usually used to harvest energy from kinetic (vibration) energy sources. The power density of electromagnetic transducers decreases as much as the effective volume of transducing material and the frequency of the energy source decrease \cite{Priya2017,Kulah2008}. Also, the fabrication and the integration of the coil and the moving magnet can come out to be difficult as the gap between the two components has to be carefully managed to prevent non-linear unpredicted responses \cite{Caruntu2001}.

@@

The hydraulic-mechanic interface can be a hydraulic cylinder (HC). It can store the energy in the frequency-up converter by actuating it at the frequency of human mastication. The amount of extractable energy from the hydraulic source will then depend on the mechanical stage behavior. Before being converted in electricity, the energy is in fact stored in a mechanical potential reservoir. During the low frequency actuation, the forces applied on each side of the HC piston are equal. Equation \ref{eq:hydraulic-mechanic_interface} models the hydraulic-mechanic interface with $x_{m}$ and $ S_{hc}$ being respectively the dynamic mass position on the $\vec{x}$ axis and the HC internal section. 
\begin{equation}
	\Delta p_{ear}S_{hc} = K_s x_m
	\label{eq:hydraulic-mechanic_interface}
\end{equation}
Therefore, the potential energy stored in a linear resonator from the ideal hydraulic source previously described can be expressed as :
\begin{equation}
	Ep_{lin} = \dfrac{K_s\ x_{m}^2}{2} = \dfrac{\Delta p_{ear} \Delta V_{ear}}{2}
	\label{eq:best_linear_soution}
\end{equation}
Thus, the linear solution can theoretically exploit/store a maximum of $50\%$ from the maximum available hydraulic energy ($\text{max}(E_{hyd})$).

Non-linear elastic potential reservoirs can also be considered. This work presents the energy extraction capability of a BR governed by a Duffing equation. It admits 2 stable equilibrium positions \mbox{$x_m=\pm x_0$}, an unstable one in \mbox{$x_m=0$} and its stiffness depends on its structural buckling level $\epsilon$ defined later in equation \ref{eq:epsilon_def}. The maximum theoretical energy in the BR for a quasi-static actuation is equal to the height of the potential energy barrier $E_{pb}$ when the mass position varies from \mbox{$x_m=\pm x_0$} to \mbox{$x_m=0$} (eq. \ref{eq:Ep_bar}). It will be a fraction $n$ of the upper limit of the available hydraulic energy (eq. \ref{eq:max_hydraulic_energy}) where $n$ has to be maximized. 
\begin{equation}
	E_{pb} = \dfrac{K x_0^2\epsilon^2}{3} = n(p_c \cdot (\Delta V_{ear})_{max})
	\label{eq:Ep_bar}
\end{equation}
We see later that it is inherently capable of exploiting $65$\% from the hydraulic energy source, which is $15$\% better than the best theoretical linear option. Section \ref{sec:SYSTEM MODELING} discusses the details of the modeling and design of the harvester that implements such a non-linear BR as it appeared to be a better solution than the linear approach. 
%/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\%
\section{HARVESTER PRESENTATION AND OPERATION \mbox{PRINCIPLE}}
\label{sec:HARVESTER PRESENTATION AND OPERATION PRINCIPLE}
%/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[!htbp]
	\centering
	\captionsetup{justification=centering}
	\includegraphics[trim={3.2cm 0cm 0cm 4.3cm},clip, width=0.7\textwidth]{figures/system_presentation.pdf}
	\caption{Schematic presentation of the frequency-up piezoelectric harvester exploiting the ear canal geometry variation} 
	\label{fig:system_presentation}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The figure \ref{fig:system_presentation} presents a schematic view of the energy harvesting system. It uses a liquid filled earplug to transmit the energy outside the ear canal. To maximize the efficiency, the system integrates a frequency-up conversion stage obtained by a BR. The latter is associated with an amplified piezoelectric generator (APG) \cite{CEDRATTECHNOLOGIES2022} made of a flextensional elastic APX4 steel \cite{AUBERT&DUVAL2022} structure containing stacked lead zirconate titanate (PZT) ceramics (fig. \ref{fig:APG}).
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%	
\begin{figure}[!htb]
	\begin{center}
		\begin{subfigure}[t]{0.5\linewidth}
			\captionsetup{justification=centering}
			\includegraphics[trim={13cm 0cm 0cm 6cm},clip,width=\linewidth]{figures/APG_schema.pdf}
			\caption{Detailed schema}
			\label{fig:APG_schema}
		\end{subfigure}
		\hfillx
		\begin{subfigure}[t]{0.21\linewidth}
			\captionsetup{justification=centering}
			\includegraphics[trim={29.5cm 0cm 0cm 8cm},clip,width=0.65\linewidth]{figures/APG_photo.pdf}
			\caption{Picture}
			\label{fig:APG_photo}
		\end{subfigure}
		\caption{The APG presentation}
		\label{fig:APG}
	\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

During a mastication cycle, the TMJ compresses the ear canal wall and expels the fluid out of the earplug. The HVs led the fluid alternatively to each HC in order to actuate the BR mass back and forth. The HA adapts both the cylinders stroke to the available earplug swept volume and the actuation force of the BR to the comfort pressure of the ear canal (fig. \ref{fig:system_presentation}).

The BR mass has two stable equilibrium positions, \mbox{$x_m = x_0$ }and \mbox{$-x_0$}. The harvester operates on two major phases. The first consists on the BR mass actuation, by a HC, from one stable equilibrium position toward the unstable position at $x_m = 0$. During the second phase the mass reaches and oscillates around the opposite stable equilibrium position until it stops, while a portion of the vibration energy is converted into electricity by the GPA. The mass will move backward when the next mastication will occur and complete the harvester cycle.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!htbp]
	\centering
	\captionsetup{justification=centering}
	\includegraphics[trim={20.5cm 0cm 0cm 11.5cm},clip, width=0.8\linewidth]{figures/HV_actuation_detail.pdf}
	\caption{Contact detail between a HV and the BR mass} 
	\label{fig:HV_actuation_detail}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

\title{Toward a novel piezoelectric energy harvester for earcanal dynamic motion exploitation, using a bistable resonator cycled by coupled hydraulic valves made of collapsed flexible tubes}

@@

%\author[critias]{Michel DEMUYNCK}
\author[critias]{Aidin DELNAVAZ}
\author[critias]{Jérémie VOIX}

\begin{abstract}
Scavenging energy from the earcanal's dynamic motion during jaw movements may be a practical way to enhance the battery autonomy of hearing aids. The main challenge consists in optimizing the amount of energy extracted while working with soft human tissues and the shape of the earcanal. This paper proposes a new harvester concept: a liquid-filled earplug which transfers energy outside the earcanal to a generator. It is composed of a hydraulic amplifier, two hydraulic cylinders that actuate a bistable resonator to raise the source frequency driving an amplified piezoelectric transducer to generate electricity. The cycling of the resonator is achieved using two innovative flexible hydraulic valves based on the buckling of flexible tubes.\\
A multiphysics-coupled model is established to determine the system operation requirements and to evaluate its theoretical performance. This model exhibits a theoretical energy conversion efficiency of 85\%. The electromechanical performance of the resonator coupled to the piezoelectric transducer as well as the hydraulic behavior of the valves are experimentally investigated. The global model is updated with the experimental data to ameliorate its predictability to further optimize the design. The energy losses are identified, which will enable the enhance the entire design proposed and improve the experimental energy conversion efficiency to 26\%.
\end{abstract}

\begin{keyword}
Energy harvesting, frequency-up, earcanal, multiphysics, hydraulic valve
\end{keyword}

@@

The growing use of wireless devices and the miniaturization of electronic circuits have led to significant progress in the energy consumption of wearable mobile devices. Energy harvesting methods have been studied to complement and enhance the power supply and autonomy of batteries. In-ear devices such as hearing aids and cochlear implants are powered by disposable cells or rechargeable batteries. The energy consumption of the best integrated devices in the literature reaches up to $17$J for a 10 hour period of daily use \cite{Scherer2019,Yip2015,Kulah2022}. Woodruff \emph{et al.} underlined that patients using hearing aids sometimes struggle to change and select batteries for their devices \cite{Woodruff2021}. Another statistical study revealed that the majority of hearing aid and cochlear implant users prefer disposable batteries for their length of autonomy, but would prefer rechargeable solutions if these were able to last the entire day \cite{PracticesAudiology2016}. Indeed, the long-term of rechargeable solutions makes them more economical and ecological compared to disposable cells. Such considerations have spurred the development of energy-harvesting systems.

The common sources of renewable energy are solar or wind but their use is unfeasible for the scale of this application. Another avenue comes from the fact that the human body generates a considerable amount of energy from various types of movements or conditions. Among them, the electrochemical energy from the inner ear \cite{Mercier2012}, kinetic energy from walking or head movements \cite{Azimi2021,Smilek2016}, thermal energy from body heat \cite{Kim2014}, and strain energy from skin deformation \cite{Jin2021} could be harvested for in-ear applications. An interesting energy source for cochlear implants or hearing aids appears to be the earcanal's mechanical deformation. Delnavaz \emph{et al.} first showed in 2012 that the jaw movements change the earcanal's geometry during mastication. Carioli \emph{et al.} then used a customized earplug molding technique with a 3D scanner to characterize the global bending movement and local compression area in the earcanal during mastication \cite{Carioli2016}. These two mechanical deformations could potentially be exploited as an energy source.

Delnavaz \emph{et al.} developed a flexible piezoelectric harvester made of a silicon earplug with a PVDF layer \cite{Delnavaz2013}. The earplug was molded to the subject's earcanal to maximize surface contact. The aim of the harvester was to exploit both the flexion and compression energies. The experimental prototype was able to extract $44$\,µJ from a mastication cycle. The soft materials facilitated the bio-integration of the harvester, but the PVDF's low electromechanicalcoupling coefficient may have been the cause of poor energy conversion efficiency. A hydro-electromagnetic energy harvester was also developed by the same researchers \cite{Delnavaz2012}. The energy was extracted using a liquid-filled earplug whose internal volume varied as a result of the radial compression of the earcanal during jaw movements. The earplug was equipped with a magnet in a water column surrounded by a coil. When the volume varied, the magnet would shift, generating energy. The generated energy experimentally amounted to 0.2µJ per mastication cycle. The prototype appeared to be less efficient than the PVDF earplug. This could be explained by the poor power generation of the electromagnetic transduction for a low frequency energy source (1.57\,Hz) and a small-scale application \cite{Kulah2008,Priya2017}. However, in this proposed model, the hydraulic transmission will enable the use of more complex and voluminous technological solutions since the energy will be made available outside, in other words, it will not be constricted by the earcanal's volume.

Based on the analysis of the literature, we propose a new solution to optimize the energy conversion efficiency and maximize the harvested energy from the earcanal's dynamic motion. The next section states the hydraulic energy source characterization and harvesting strategy that was adopted. Section \ref{sec:HARVESTER PRESENTATION AND OPERATION PRINCIPLE} presents the harvester and describes its operation principle. Section \ref{sec:SYSTEM MODELING} details the global system modelling and section \ref{sec:NUMERICAL MODEL AND SIMULATIONS} presents the numerical simulation of the theoretical behavior of the harvester. Then section \ref{sec:EXPERIMENTAL CHARACTERIZATIONS OF THE ELECTROMECHANICAL CONVERTER} exposes the experimental validation performed on the electromechanical converter stage and section \ref{sec:EXPERIMENTAL APPROACH FOR THE HV DESIGN} shows the experimental approach for the design of the hydraulic valves for the system cycling. Section \ref{sec:MODEL RECALIBRATION WITH EXPERIMENTAL DATA} presents the experimentally adjusted global harvester model and discusses this work's major contributions and the potential of the harvester. This is followed by the conclusion.

%/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\%
\section{ENERGY SOURCE AND HARVESTING STRATEGY}
\label{sec:THE ENERGY SOURCE AND HARVESTING STRATEGY}
%/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\%
    %///////////////////////////////////////////// 
	\subsection{The energy source}	
	\label{The energy source}
    %/////////////////////////////////////////////
Extracting energy from the mechanical deformations of human tissue requires soft flexible materials so that the earcanal deflection energy can be harvested without causing discomfort. However, flexible transducers have a low energy-conversion efficiency because part of the extracted energy is dissipated in the elastic non-active material. Additionally, the low frequency of mastication makes it difficult to use resonant structures that generally admit better conversion efficiencies than low frequency solutions \cite{Ashraf2011}. Thus, the traditional transducing methods are unsuitable for this application.

To date, the best solution proposed in the literature consists of using soft materials, such as the piezoelectric PVDF earplug previously introduced \cite{Delnavaz2013}. In fact, the low mechanical impedance of the soft tissue limits the use of more effective rigid piezoelectric materials.

The hydraulic extraction strategy proposed in \cite{Delnavaz2012} is less effective than the piezoelectric earplug but energy can be transmitted outside the earcanal which could benefit from larger effective space. The latter is essential to be able to implement an impedance adaptation stage and more effective transducing methods that could not otherwise fit directly inside the earcanal's limited volume. In fact, Marin \emph{et al.} demonstrated that the power generated by piezoelectric or electromagnetic transducers is proportionally greater as the effective volume of the transducing material is greater \cite{Marin2011}. Thus, the present work is based on the exploitation of the earcanal deformation using a liquid-filled earplug. As an energy source, it can be assimilated to a micro pump characterized by its volume variation $V_{ear}$ and pressure variation $p_{ear}$. In order to optimize the energy extraction, the earplug must fit the earcanal wall snugly. Turcot \emph{et al.} demonstrated that the initial pressurization must reach a minimum of $14$kPa to achieve a good fit \cite{TURCOT2011}. Bouchard-Roy \emph{et al.} studied the dynamic pressure variation inside the earplug for different initial pressurization levels \cite{Bouchard-Roy2020}. A maximum of $\text{max}(\Delta p_{ear})=12$kPa of dynamic pressure variation, with a $28$kPa of initial pressurization, has been noted during the mastication for seven subjects. Besides, the temporal evolution of the volume variation $\Delta V_{ear}$ has been estimated in \cite{Delnavaz2012} and the result for a mastication cycle is presented on figure \ref{fig:deltaV_ear}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!htbp]
	\centering
	\captionsetup{justification=centering}
	\includegraphics[trim={20.5cm 0cm 0cm 10.8cm},clip, width=0.4\textwidth]{figures/deltaV_ear.pdf}
	\caption{Earcanal volume variation for one mastication cycle}
	\label{fig:deltaV_ear}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The earplug operation characteristic behaviour has not been studied yet. Thus, the evolution of the volume variation $\Delta V_{ear}$ is not well known as a function of the pressure variation $\Delta p_{ear}$. Then, considering the pressure and volume variation levels from \cite{Delnavaz2012} and \cite{Bouchard-Roy2020}, the equation \ref{eq:max_hydraulic_energy} approximates an upper limit for the extractable hydraulic energy with the maximum values of $\Delta p_{ear}$ et $\Delta V_{ear}$ that have been recorded, i.e. $p_c=12$\,kPa and $(\Delta V_{ear})_{max}=60$\,mm$^3$. $p_c$ stands for the comfort pressure by assuming that the user could feel discomfort if $\Delta p_{ear}>p_c$. The energy estimation presumes that the earplug is a perfect flow rate source, which is a theoretical hypothesis leading to $0.9$\,mJ available from one mastication. Based on 2200 mastication cycles per day, according to \cite{Goll2011}, there would be $2$\,mJ available per day, per ear. This amount of energy could theoretically enhance supply $23.5$\,\% of the $17$\,J of needed for the state of the art of cochlear implants \cite{Kulah2022}. Further research in both domains related to hearing aids and energy harvesting could lead to fully autonomous devices in the future.
\begin{align}
	\text{max}(E_{hyd}) = p_c \cdot (\Delta V_{ear})_{max} = 0.9\text{\,mJ}
	\label{eq:max_hydraulic_energy}
\end{align}

    %///////////////////////////////////////////// 
	\subsection{The harvesting strategy}	
	\label{The harvesting strategy}
    %/////////////////////////////////////////////
Electromagnetic transducers are usually used to harvest energy from kinetic (vibration) energy sources. In fact, the power density of electromagnetic transducers decreases as much as the effective volume of transducing material and the frequency of the energy sources decreases \cite{Priya2017,Kulah2008}. Also, the fabrication and the integration of the coil and the moving magnet can come out to be difficult as the gap between the two components has to be carefully managed to prevent non-linear unpredicted responses \cite{Caruntu2001}.

@@

The hydraulic-mechanic interface can be a hydraulic cylinder (HC). It can store the energy in the frequency-up converter by actuating it at the frequency of human mastication. The amount of extractable energy from the hydraulic source will then depend on the mechanical stage behavior. Before being converted in electricity, the energy is in fact stored in a mechanical potential reservoir. During the low frequency actuation, the forces applied on each side of the HC piston are equal. Equation \ref{eq:hydraulic-mechanic_interface} models the hydraulic-mechanic interface with $x_{m}$ and $ S_{hc}$ being respectively the dynamic mass position on the $\vec{x}$ axis and the HC internal section.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{equation}
	\Delta p_{ear}\ S_{hc} = K_{s}\ x_{m}
	\label{eq:hydraulic-mechanic_interface}
\end{equation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Therefore, the potential energy stored in a linear resonator from the ideal hydraulic source previously described can be expressed as follows :
\begin{equation}
	Ep_{lin} = \dfrac{K_s\ x_{m}^2}{2} = \dfrac{\Delta p_{ear} \Delta V_{ear}}{2}
	\label{eq:best_linear_soution}
\end{equation}
Thus, the linear solution can theoretically exploit/store a maximum of $50\,\%$ from the maximum available hydraulic energy ($\text{max}(E_{hyd})$).

Non-linear elastic potential reservoirs can also be considered. This work presents the energy extraction capability of a BR governed by a Duffing equation. It admits 2 stable equilibrium positions \mbox{$x_m=\pm x_0$}, an unstable one in \mbox{$x_m=0$} and its stiffness depends on its structural buckling level $\epsilon$ defined later in equation \ref{eq:epsilon_def}. The maximum theoretical energy in the BR for a quasi-static actuation is equal to the height of the potential energy barrier $E_{pb}$ when the mass position varies from \mbox{$x_m=\pm x_0$} to \mbox{$x_m=0$} (eq. \ref{eq:Ep_bar}). It will be a fraction $n$ of the upper limit of the available hydraulic energy (eq. \ref{eq:max_hydraulic_energy}) where $n$ has to be maximized. 
\begin{equation}
	E_{pb} = \dfrac{K x_0^2\epsilon^2}{3} = n(p_c \cdot (\Delta V_{ear})_{max})
	\label{eq:Ep_bar}
\end{equation}
We see later that it is inherently capable of exploiting $65$\,\% from the hydraulic energy source, which is $15$\,\% better than the best theoretical linear option. Section \ref{sec:SYSTEM MODELING} discusses the details of the modeling and design of the harvester that implements such a non-linear BR as it appeared to be a better solution than the linear approach. 
%/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\%
\section{HARVESTER PRESENTATION AND OPERATION \mbox{PRINCIPLE}}
\label{sec:HARVESTER PRESENTATION AND OPERATION PRINCIPLE}
%/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[!htbp]
	\centering
	\captionsetup{justification=centering}
	\includegraphics[trim={3.2cm 0cm 0cm 4.3cm},clip, width=0.7\textwidth]{figures/system_presentation.pdf}
	\caption{Schematic presentation of the frequency-up piezoelectric harvester exploiting the earcanal geometry variation} 
	\label{fig:system_presentation}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The figure \ref{fig:system_presentation} presents a schematic view of the energy harvesting system. It uses a liquid filled earplug to transmit the energy outside the earcanal. To maximize the efficiency, the system integrates a frequency-up conversion stage obtained by a BR. The latter is associated with an amplified piezoelectric generator (APG) \cite{CEDRATTECHNOLOGIES2022} made of a flextensional elastic APX4 steel \cite{AUBERT&DUVAL2022} structure containing stacked lead zirconate titanate (PZT) ceramics (fig. \ref{fig:APG}).
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%	
\begin{figure}[!htb]
	\begin{center}
		\begin{subfigure}[t]{0.5\linewidth}
			\captionsetup{justification=centering}
			\includegraphics[trim={13cm 0cm 0cm 6cm},clip,width=\linewidth]{figures/APG_schema.pdf}
			\caption{Detailed schema}
			\label{fig:APG_schema}
		\end{subfigure}
		\hfillx
		\begin{subfigure}[t]{0.21\linewidth}
			\captionsetup{justification=centering}
			\includegraphics[trim={29.5cm 0cm 0cm 8cm},clip,width=0.65\linewidth]{figures/APG_photo.pdf}
			\caption{Picture}
			\label{fig:APG_photo}
		\end{subfigure}
		\caption{The APG presentation}
		\label{fig:APG}
	\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

During a mastication cycle, the TMJ compresses the earcanal wall and expels the fluid out of the earplug. The HVs led the fluid alternatively to each HC in order to actuate the BR mass back and forth. The HA adapts both the cylinders stroke to the available earplug swept volume and the actuation force of the BR to the comfort pressure of the earcanal (fig. \ref{fig:system_presentation}).

The BR mass has two stable equilibrium positions, \mbox{$x_m = x_0$ }and \mbox{$-x_0$}. The harvester operates on two major phases. The first consists on the BR mass actuation, by a HC, from one stable equilibrium position toward the unstable position at $x_m = 0$. During the second phase the mass reaches and oscillates around the opposite stable equilibrium position until it stops, while a portion of the vibration energy is converted into electricity by the GPA. The mass will move backward when the next mastication occurs and complete the harvester cycle.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!htbp]
	\centering
	\captionsetup{justification=centering}
	\includegraphics[trim={20.5cm 0cm 0cm 11.5cm},clip, width=0.8\linewidth]{figures/HV_actuation_detail.pdf}
	\caption{Details of the contact between a HV and the BR mass} 
	\label{fig:HV_actuation_detail}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
